---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mxnorm: An R package to normalize multiplexed imaging data.

<!-- badges: start -->
<!-- badges: end -->

A package designed to handle multiplexed imaging data in R, implementing normalization methods and quality metrics detailed in our paper [here](https://doi.org/10.1101/2021.07.16.452359).

# Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ColemanRHarris/mxnorm")
```
# Analysis Example

This is a basic example using the `mx_sample` dataset, which is simulated data to demonstrate the package's functionality.

```{r sample}
library(mxnorm)
head(mx_sample)
```

## `mx_dataset` objects

How to build the `mx_dataset` object with `mx_sample` data in the `mxnorm` package:

```{r example}
mx_dataset = mx_dataset(data=mx_sample,
                        slide_id="slide_id",
                        image_id="image_id",
                        marker_cols=c("marker1_vals","marker2_vals","marker3_vals"),
                        metadata_cols=c("metadata1_vals"))
```

The structure of the now built `mx_dataset` object:

```{r ex str}
str(mx_dataset)
```

## Normalization with `mx_normalize()`

And now we can normalize this data using the `mx_normalize()` function:
```{r norm}
mx_norm = mx_normalize(mx_data = mx_dataset,
                       scale = "log10",
                       method="None")
```

The `mx_dataset` object has normalized data in the following form:

```{r norm data}
head(mx_norm$norm_data)
```

And the `mx_dataset` object now includes the following attributes:

```{r norm str}
str(mx_norm)
```

## Otsu misclassification metrics with `otsu_misclass()`

Using the above normalized data, we can run misclassification metrics to determine how well our normalization method performs:
```{r otsu}
mx_otsu = otsu_misclass(mx_norm,
                        table="both",
                        threshold_override = NULL,
                        plot_out = FALSE)
```

This adds an Otsu misclassification table to the `mx_dataset` object in the following form:
```{r}
head(mx_otsu$otsu_data)
```

And the `mx_dataset` object now includes the following attributes:

```{r otsu str}
str(mx_otsu)
```

# Visualizations

We can also begin to visualize these results using some of `mxnorm`'s plotting features built using `ggplot2`. 

First, we can visualize the densities of the marker values as follows:

```{r mx dens}
plot_mx_density(mx_otsu)
```

We can also visualize the results of the Otsu misclassification analysis stratified by slide and marker:

```{r mx var}
plot_mx_variance(mx_otsu)
```
